Visualizing Algebraic Surfaces

### 2016-04-14 project/article/src/article.tex
moved over projective geom stuff

>>>>

    Quite informally, projective space formalizes the notion of parallell lines
    intersecting at infinity. In order to get a better understanding of the
    general projective space $\proj^n$ we will first construct the \emph{real
    projective plane}\index{real projective plane} $\proj^2(\R)$.

\subsection{The real projective plane}

    According to \cite{Wik16} there are three equivalent definitions:
    \begin{enumerate}[I.]
        \item\label{def:one} The set of all lines in $\R^3$ passing through the origin $\left( 0,
            0, 0 \right)$. Every such line meets the \emph{sphere} of radius one
            centered in the origin exactly twice, say in $P = (x, y, z)$ and its
            antipodal point $(-x, -y, -z)$.
        \item\label{def:two} The points on the sphere $S^2$, where every point $P$ and its
            antipodal points are not distinguished. For example, the point $(1, 0,
            0)$ is identified with the point $(-1, 0, 0)$.
        \item The set of equivalence classes of $\R^3 \setminus (0, 0, 0)$ where
            two points $P$ and $P^\prime$ are considered equivalent if and only if
            there is a positive $\lambda \in \R$ such that $P = \lambda P^\prime$.
    \end{enumerate}

    The last definition is the one seemingly most commonly used, and is therefore
    the one we will employ here. Note that the elements in the real projective
    plane are equivalence classes of points in $\R^3$. We denote an element in
    $\proj^2(\R)$ as $\projp{x}{y}{z}$. These elements are commonly referred to as
    \emph{homogenous coordinates}\index{homogenous coordinates}. We state this
    formally in a definition.

    \begin{defn}[The real projective plane]\index{$\proj^2(\R)$}
        Let $\sim$ be the equivalence relation on $\R^3$ defined by
        \begin{equation}
            \notag
            (x, y, z) \sim (x', y', z') \iff (x, y, z) = \lambda(x, y, z),
        \end{equation}
        where $\lambda$ is some positive real number. Then we denote the
        equivalence class of $(x, y, z)$ as $\projp{x}{y}{z}$ and we define
        \emph{the real projective plane} as the set
        \begin{equation}
            \notag
            \proj^2(\R) = \left\{ \projp{x}{y}{z} \mid (x, y, z) \in \R^3 \setminus
            (0, 0, 0) \right\}.
        \end{equation}
    \end{defn}

    We can also look at a subset of $\proj^2(\R)$, namely the set of points
    $\projp{x}{y}{z}$ where $z \neq 0$. Remembering the identification from the
    above definition, these are equivalent to $\projp{\frac{x}{z}}{\frac{y}{z}}{1}$
    with $\lambda = 1 / z$. If we consider definition \ref{def:two} from above,
    this corresponds to the set of points on the northern hemisphere of $S^2$, not
    including the circle of intersection in the $(x,y)$-plane. Similarly, we can
    look at the set of points $\projp{x}{y}{0}$. This corresponds, in $S^2$, to the
    circle of intersection in the $(x, y)$-plane.

    We can now take the notion of the projective plane and generalize to $n$
    dimensions.
    \subsection{Projective Space}
    \label{sub:projective_space}

    The general definition of projective $n$-space is completely anologous to the
    definition of the projective plane, the set of lines in $\R^{n+1}$ passing
    through the origin, however we include a formal definition for completeness.

    \begin{defn}[The real projective space of dimension $n$]\index{projective
        space}\index{$\proj^n(\R)$}

        Let $\sim$ be an equivalence relation on $\R^{n+1}\setminus \left( 0,
        \ldots, 0 \right)$ defined by
        \begin{equation}
            \notag
            \left( x_1, \ldots, x_{n+1} \right) \sim \left( x_1', \ldots, x_{n+1}' \right) 
            \iff \left( x_1, \ldots, x_{n+1} \right) = \lambda \left( x_1', \ldots, x_{n+1}' \right)
        \end{equation} 
        where $\lambda$ is some positive real number. We then denote the
        equivalence class of $\left( x_1, \ldots, x_{n+1} \right)$ by
        $\projp{x_1}{\ldots}{x_{n+1}}$ and define \emph{the real projective space
        of dimension $n$} as the set
        \begin{equation}
            \notag
            \proj^n(\R) = \left\{ \projp{x_1}{\ldots}{x_{n+1}} \mid \left( x_1, \ldots, x_{n+1} \right) 
            \in \R^{n+1} \setminus \left( 0, \ldots, 0 \right) \right\}.
        \end{equation}
    \end{defn}
